### 2024-09-18 article_codes.qmd
add touche = TRUE when extracting values from a SpatRaster.
Add a table with number of buildings ein each WUI classification (standard WUI)

# Summary by county
tbl_wui_county <- terra::extract(wui_r, counties) |>  
  pivot_longer(cols = -ID, 
               names_to = "name",
               values_to = "value") |> 
  group_by(ID, name, value) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(ID, name) |> 
  mutate(total = sum(count), 
         pct = 100*count/total) |> 
  ungroup() |> 
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |> 
  mutate(pct_wui = pct_Interface + pct_Intermix)

@@

# Number of buildings by county  ----
county_build_count <- build_counts |> 
  # Number of buildings per county
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |> 
  # sf object
  st_as_sf() %>% 
  # Area polygons
  mutate(area = units::set_units(st_area(.), "km2"))  |> 
  # Building density by county
  mutate(build_den = sum/area)

@@

# Function for getting vegetation area by counties 
f_pct_veg <- function(df) {
  df |> # sf object with vegetation data  
    # Number of cells
    extract(y = vect(counties),
            fun = "sum",
            na.rm = TRUE,
            bind = TRUE) |>  
    st_as_sf() |> 
    # Area (number cells x area cells)
    mutate(
      area_km = wild_veg * ((res(df))[1]/1000 * (res(df ))[2]/1000)
    ) |> 
    # Get only values 
    st_drop_geometry() |> 
    select(NUTS_NAME, CNTR_CODE, area_km)
}

@@

@@

# Link wildfires with building density (build_focal) 
wf_build_density <- extract(build_focal, wf) |>
  # Assign 0 to NAs --> No buildings
  mutate(focal_sum = ifelse(is.na(focal_sum), 0, focal_sum)) |> 
  # Building density is the number of buildings / buffer area 
  # buffer area = circular moving window of 500 m
  mutate(bd = focal_sum / area_buffer)

>>>>

# Summary by county
tbl_wui_county <- terra::extract(wui_r, counties, touches = TRUE) |>  
  pivot_longer(cols = -ID, 
               names_to = "name",
               values_to = "value") |> 
  group_by(ID, name, value) |> 
  summarise(count = n()) |> 
  ungroup() |> 
  group_by(ID, name) |> 
  mutate(total = sum(count), 
         pct = 100*count/total) |> 
  ungroup() |> 
  pivot_wider(names_from = "value",
              values_from = c("count", "pct")) |> 
  mutate(pct_wui = pct_Interface + pct_Intermix)

@@

# Number of buildings by county  ----
county_build_count <- build_counts |> 
  # Number of buildings per county
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          touches = TRUE,
          bind = TRUE) |> 
  # sf object
  st_as_sf() %>% 
  # Area polygons
  mutate(area = units::set_units(st_area(.), "km2"))  |> 
  # Building density by county
  mutate(build_den = sum/area)

@@

# Function for getting vegetation area by counties 
f_pct_veg <- function(df) {
  df |> # sf object with vegetation data  
    # Number of cells
    extract(y = vect(counties),
            fun = "sum",
            na.rm = TRUE,
            touches = TRUE,
            bind = TRUE) |>  
    st_as_sf() |> 
    # Area (number cells x area cells)
    mutate(
      area_km = wild_veg * ((res(df))[1]/1000 * (res(df ))[2]/1000)
    ) |> 
    # Get only values 
    st_drop_geometry() |> 
    select(NUTS_NAME, CNTR_CODE, area_km)
}

@@

Number of buildings in WUI

```{r}
#| label: tbl-buildings-wui
#| tbl-cap: Number of buildings in each WUI classification

c(build_counts, wui_r) |> 
  as_tibble() |>   
  mutate(sum = ifelse(is.na(sum), 0, sum)) |> 
  drop_na() |> 
  pivot_longer(col = -sum, names_to = "name") |> 
  group_by(name, value) |>
  summarise(buildings = sum(sum)) |> 
  ungroup() |> 
  pivot_wider(names_from = name, 
              values_from = buildings) |> 
  gt(rowname_col =  "value") |> 
  grand_summary_rows(
    fns = list(Total = ~sum(.))
  ) |> 
   cols_label(
    corine = md("**corine**"),
    effis = md("**effis**"),
    firEUrisk ~ md("**firEUrisk**"),
    national ~ md("**national**")
  )

@@

# Link wildfires with building density (build_focal) 
wf_build_density <- extract(build_focal, wf, touches = TRUE) |>
  # Assign 0 to NAs --> No buildings
  mutate(focal_sum = ifelse(is.na(focal_sum), 0, focal_sum)) |> 
  # Building density is the number of buildings / buffer area 
  # buffer area = circular moving window of 500 m
  mutate(bd = focal_sum / area_buffer)
